---
title: "Sci Reports mixed genotypes resist disease"
author: "Anya Brown"
date: "11/04/22"
output: 
  html_document: 
    fig_height: 10
    fig_width: 10
---
Coral health statuses were monitored for disease in the CCMI nursery 
Dataset is located in Dryad: 
Brown, Anya et al. (2022), CCMI nursery coral disease 2019, Dryad, Dataset, https://doi.org/10.25338/B8F643

```{r setup, warning = FALSE, message = FALSE, results = FALSE}
library(tidyr)
library(dplyr)
library(reshape2)
library(plyr)
library(ggplot2)
library(Rmisc)
library(stringr)
library(lme4)
library(lmerTest)
library(cowplot)
library(patchwork)
library(viridis)
library(glmmTMB)
library(DHARMa)
#dataset in Dryad 

```

Data set up
```{r setup data, warning=FALSE, message = FALSE}
dis_1 <-read.csv("~/Dropbox/POST DOC/PostDoc Project/Spring 2019/Research/Research/Disease_deep nursery/Frame Surveys/Data/FrameSurvey_Dec2019_wideformat_edit2.csv")

setwd("~/Dropbox/POST DOC/PostDoc Project/Spring 2019/Research/Research/Disease_deep nursery/Frame Surveys/Data/")
#if goes from dead to missing, change back to dead 
library(data.table)
```

Make sure data in columns are consistently named 
```{r data management}
dis <- reshape2::melt(dis_1, id= c("Frame", "Column","Row","Genotype")) #value.name ="Coral_stat")
colnames(dis)[5] <- "Date"
colnames(dis)[6] <- "Coral_stat"

#remove X from the dates
dis$Date <-gsub("X","",dis$Date)

#create unique IDs for each coral based on frame number and location on a frame
dis$ID <- paste(dis$Frame, dis$Column, dis$Row)

#remove Xs from data
#dis$Date <- str_remove(dis$Date,"X")
#remove leading or trailing white spaces
dis$Coral_stat <- trimws(dis$Coral_stat)

#Replace P with H
p <- which(dis$Coral_stat == "P")
dis[p,"Coral_stat"] <- "H"


#Replace XS with X
s <- which(dis$Coral_stat == "XS")
dis[s,"Coral_stat"] <- "X"

#Replace HR with H (Healty recovred should be just health)
h <- which(dis$Coral_stat == "HR")
dis[h,"Coral_stat"] <- "H"

#replace "0" with O for missing
o <- which(dis$Coral_stat == "0")
dis[o,"Coral_stat"] <- "O"

#replace OD with Dead
d <- which(dis$Coral_stat == "OD")
dis[d,"Coral_stat"] <- "D"

#replace R for Recovered with Healthy
e <- which(dis$Coral_stat == "R")
dis[e,"Coral_stat"] <- "H"

#replace diseased and recovered (XR) with X (disease)
f <- which(dis$Coral_stat == "XR")
dis[f,"Coral_stat"] <- "X"

#replace recovered with disease with just disease
g <- which(dis$Coral_stat == "RX")
dis[g,"Coral_stat"] <- "X"

#remove data from May 16 - not trustworthy
dis <- dis[which(dis$Date != "2019.05.16"),]
#remove missing corals (never present) and where there is no data
dis <- dis[which(dis$Coral_stat != "O" & dis$Coral_stat != "NoData"),]
```


Genotype counts
```{r genotype counts}
counts_dis_g <- plyr::ddply(na.omit(dis), c("Date", "Genotype","Frame","Coral_stat"), summarize,
      count = length(Coral_stat))

```

Including zeroes for other health categories  
```{r include zeroes}
#put all data longwise 
counts_dis_g <- as.data.table(counts_dis_g)
dim(counts_dis_g)
counts_cast <- dcast.data.table(counts_dis_g, Date+Genotype+Frame~Coral_stat, value.var = "count")

counts_cast <- counts_cast %>% mutate_all(~replace(., is.na(.), 0))
dim(counts_cast)
```

Melt data 
```{r melt data}
counts_cast_melt <- melt(data = counts_cast, id.vars = c("Date","Frame","Genotype"))
colnames(counts_cast_melt)[4] <- "Coral_stat"
colnames(counts_cast_melt)[5] <- "count"

```
Labeling single and mixed genotype frames
```{r single mixed genotype label}
library(data.table)
frame <- data.table(unique(counts_cast_melt$Frame))
colnames(frame)<- "Frame"
frame$sing_mix <- ifelse(frame$Frame > 17, "mixed","single")
s <- which(frame$Frame == "1207")
frame[s,"sing_mix"] <- "single"
w <- which(frame$Frame == "1279")
frame[w,"sing_mix"] <- "single"


```

Counts across all of the frames and genotypes
```{r setting up data}
#count number of corals in a health category - this sums the number of corals over all health categories on a frame to get the total number of corals on a frame

counts_fr <- ddply(counts_cast_melt, c("Date", "Frame"), summarize,       countf = sum(count))

#for each frame, wheter it's a single or mixed genotype
singmixf <- merge(frame, counts_fr)

#total number of corals on single or mixed frames at each date
total.per <- ddply(singmixf, c("Date","sing_mix"), summarize, 
      total = sum(countf))

#total number of frames 
total.f <- ddply(singmixf, c("Date","sing_mix"), summarize, 
      total = length(Frame))
#12 mixed frames
#18 single frames 

#432 single corals
#218 mixed
```

Calculation of proportion of health statuses on a frame including zeroes
```{r health cat proportion}

counts_all_singmix <- merge(counts_cast_melt, singmixf, by = c("Frame", "Date"))
dim(counts_all_singmix)


counts_all_of_frame <- ddply(counts_all_singmix, c("Date", "Frame","sing_mix","countf","Coral_stat"), summarize,       
                  count_all = sum(count))
dim(counts_all_of_frame)
#average over all frames
sum_all_counts <- Rmisc::summarySE(counts_all_of_frame, measurevar = "count_all", groupvars = c("Date","Coral_stat","sing_mix"))


counts_all_of_frame$prop <- counts_all_of_frame$count_all/counts_all_of_frame$countf

#average of proportion in each health category per frame 
sum_all <- Rmisc::summarySE(counts_all_of_frame, measurevar = "prop", groupvars = c("Date","Coral_stat","sing_mix"))


```


Figure 2: Plot of Disease for all corals across the nursery 

```{r Figure 2 - Single vs mixed All}

sum_all2 <- subset(sum_all, Date != "2019.02.01")
sum_all2$Date2 <- gsub("[.]", "/", sum_all2$Date)
sum_all2$Date3 <- as.Date(sum_all2$Date2)

#### Used in Figure 2 below ####
sing_mix_all  <- ggplot(sum_all2[which(sum_all2$Coral_stat=="X"),], aes(x = Date3, y = prop, group =sing_mix)) + geom_point(aes(fill = sing_mix), pch = 21, color = "black",size = 3, alpha= 0.75, position = position_dodge(width = 0.5)) + geom_errorbar(aes(ymax = prop+se, ymin = prop-se), position = position_dodge(width = 0.5)) + ylab("Disease Prevalance") + theme_bw() + scale_fill_manual(values = c("darkorchid4","gold"), labels = c("mixed", "single"), name = "Diversity") + theme(axis.text.x = element_text(angle = 90),panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "bottom") + xlab("Date")

```


Disease-only data 
```{r disease and dead data}
disease <- counts_all_of_frame[which(counts_all_of_frame$Coral_stat == "X"),]
dim(disease)
dead <- counts_all_of_frame[which(counts_all_of_frame$Coral_stat == "D"),]


total.disease.counts <- ddply(disease, c("Date","sing_mix"), summarize, 
sum.disease = sum(count_all), sum.all = sum(countf))

total.dead.counts <- ddply(dead, c("Date","sing_mix"), summarize, sum.disease = sum(count_all), sum.all = sum(countf))


#create factors and turn dates into date data type
disease$Frame <- as.factor(disease$Frame)
disease$Date2 <- gsub("[.]", "/", disease$Date)
disease$Date3 <- as.Date(disease$Date2)
#disease$Date2 <- as.numeric(disease$Date)

#includind a polynomial
x2 <-poly(disease$Date3,2)
```


Model 1 analysis (all data)
```{r set up for Model 1 analysis}
disease <- counts_all_of_frame[which(counts_all_of_frame$Coral_stat == "X"),]
disease$Frame <- as.factor(disease$Frame)
disease$Date2 <- gsub("[.]", "/", disease$Date)
disease$Date3 <- as.Date(disease$Date2)

#include density
Frame_area <- read.csv("~/Dropbox/POST DOC/PostDoc Project/Spring 2019/Research/Research/Disease_deep nursery/Frame Surveys/Data/Frame_area.csv")

#dim(disease)
#dim(disease_dens)
disease_dens <- merge(disease, Frame_area, by = "Frame")
disease_dens$density <- disease_dens$countf/disease_dens$Area
disease_dens$density_rescale <- reshape::rescaler(to = c(0,1), x = disease_dens$density)
```

Model 1: Binomial analysis
```{r binomial Model 1}
x2 <-poly(disease_dens$Date3,2)

library(glmmTMB)
glmmtb_alldat <- glmmTMB(cbind(count_all, countf-count_all) ~ sing_mix*x2+density + (1|Frame), family = binomial, data= disease_dens)

```

Check Model 1 residuals
```{r binomial residual glmtbb}
library(DHARMa, quietly = TRUE)
glmtb_resid1 <- simulateResiduals(glmmtb_alldat)
testOutliers(glmtb_resid1, type = "bootstrap")
plot(glmtb_resid1)
```

Summary of Model 1 as a table
```{r table for glmmtb}

summary(glmmtb_alldat)
knitr::kable(car::Anova(glmmtb_alldat))
```

Density information
```{r counts information, results = FALSE}
disease_dens %>%
   dplyr::group_by(Date,sing_mix)%>%
  dplyr::summarize(total = sum(count_all))
```

Figure 1 data summary
```{r Figure 1: data summary}
sum_all2 <- Rmisc::summarySE(counts_all_of_frame, measurevar = "prop", groupvars = c("Date","Coral_stat"))

sum_all2$Date2 <- gsub("[.]", "/", sum_all2$Date)
sum_all2$Date3 <- as.Date(sum_all2$Date2)


```

Figure 1 Plot 
```{r Figure 1 plot}
#### Figure 1 ####

ggplot(sum_all2[which(sum_all2$Date!="2019.02.01"),], aes(x = Date3, y = prop, group = Coral_stat)) + geom_point(aes(fill = Coral_stat),color = "black", pch = 21, size = 3) + theme_bw() + geom_errorbar(aes(ymax = prop+se, ymin = prop-se)) + ylab("Proportion of health status over time")  + theme(axis.text.x = element_text(angle = 90), panel.grid.minor = element_blank(), panel.grid.major = element_blank()) + scale_fill_manual(values = c("black","orange", "gray"), name = "Coral Health", labels = c("Dead","Healthy", "Disease")) + xlab("Date")

```

Figure S2: Health statuses across all frames set up
```{r Figure S2: Health statuses across all frames}
#### Supplement - all Frames
sum_all3 <- Rmisc::summarySE(counts_all_of_frame, measurevar = "prop", groupvars = c("Date","Coral_stat", "Frame", "sing_mix"),na.rm = T)

sum_all3$Date2 <- gsub("[.]", "/", sum_all3$Date)
sum_all3$Date3 <- as.Date(sum_all3$Date2)

sum_all3$Frame <- factor(sum_all3$Frame, levels = c("1235","1231","14","9","1237","7","5","1221","1219","12","11","1251","1201","1217","13","10","1209","6","4","3","8","1279","1207","16","15","1107","1","2","1295","1296"))
```

Figure S2: Health statuses across all frames
```{r Figure S2 Supplement Plot: All frames, out.width="50%"}
ggplot(sum_all3[which(sum_all3$Date!="2019.02.01"),], aes(x = Date3, y = prop, group = Coral_stat)) + theme_bw() + geom_errorbar(aes(ymax = prop+se, ymin = prop-se)) + ylab("Proportion of health status over time")  + theme(text = element_text(size = 20), axis.text.x = element_text(size = 16, angle = 90), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_rect(color = "black", fill ="white")) + scale_color_manual(values = c("black","orange", "gray"), name = "Coral Health",labels = c("Dead","Healthy","Disease")) + xlab("Date") + facet_wrap(~Frame+sing_mix) + geom_line(aes(color = Coral_stat), size = 1)+ geom_point(aes(color = Coral_stat), size = 1)

```


Including genotype information 
calculates the proportion of disease within a genotype on a frame
```{r proportions including genotypes - genotypes on single and mixed frames}
counts_cast_melt <- melt(data = counts_cast, id.vars = c("Date","Frame","Genotype"))

colnames(counts_cast_melt)[4] <- "Coral_stat"
colnames(counts_cast_melt)[5] <- "count"

counts_cast_melt <- subset(counts_cast_melt, Coral_stat == "X" | Coral_stat == "H" | Coral_stat == "D")
dim(counts_cast_melt)
#Blue
B <- counts_cast_melt[which(counts_cast_melt$Genotype == "B"),]
counts_fr <- ddply(B, c("Date", "Frame"), summarize,      
                   countf = sum(count))
Bf <- merge(B, counts_fr, by = c("Date","Frame"))

Bf$prop <- Bf$count/Bf$countf            

genotypefunc <- function(Gen){
   Gent <- counts_cast_melt[which(counts_cast_melt$Genotype == Gen),]
   counts_fr <- ddply(Gent, c("Date", "Frame"), summarize,      
                   countf = sum(count))
Gentf <- merge(Gent, counts_fr, by = c("Date","Frame"))
Gentf$prop <- Gentf$count/Gentf$countf            
Gentf <- as.data.frame(Gentf)
}

Rf <- genotypefunc(Gen = "R")
dim(Rf)
Kf <- genotypefunc(Gen = "K")
dim(Kf)
Gf <- genotypefunc(Gen = "G")
dim(Gf)
Yf <- genotypefunc(Gen = "Y")
dim(Yf)

#combine together
GenRYGBK <- rbind(Rf, Kf, Bf, Gf, Yf)
dim(GenRYGBK)
```

Summarizing the data with genotypes 
```{r proportions including genotypes}
#merge with single mix

colnames(singmixf)[4] <- "count_all"
GenRYGBK_sm <- merge(singmixf, GenRYGBK, by = c("Date","Frame"))
#dimension check
dim(singmixf)
dim(GenRYGBK)
dim(GenRYGBK_sm) #summarize - just B, K, R, Y, G


sumgen1 <- Rmisc::summarySE(GenRYGBK_sm, measurevar = c("prop"), groupvars = c("Date","Genotype","sing_mix", "Coral_stat"))

sum_onlygrybk <- Rmisc::summarySE(GenRYGBK_sm, measurevar = c("prop"), groupvars = c("Date","sing_mix", "Coral_stat"))

```


Model 2: Including genotypes in understanding disease prevalence
```{r Model 2: disease with only genotypes on both frames}
disease2 <- GenRYGBK_sm[which(GenRYGBK_sm$Coral_stat == "X"),]
disease2$Frame <- as.factor(disease2$Frame)
disease2$sing_mix <- as.factor(disease2$sing_mix)
disease2$Genotype <- as.factor(disease2$Genotype)

disease2$Date2 <- gsub("[.]", "/", disease2$Date)
disease2$Date3 <- as.Date(disease2$Date2)
#between genotype and single mixed
disease2_dens <- merge(Frame_area, disease2, by = "Frame")
dim(disease2)
dim(disease2_dens)
disease2_dens$density <- disease2_dens$countf/disease2_dens$Area
library(stats)
x2 <-poly(disease2_dens$Date3,2)
```


```{r Model 2: binomial  including genotypes glmmTMB}
library("glmmTMB")

glmmtb_all <- 
glmmTMB(cbind(count, countf-count) ~ sing_mix*x2*Genotype + density + (1|Frame), family = binomial, data= disease2_dens)

```


```{r Model 2 binomial  including genotypes glmmTMB summary}
summary(glmmtb_all)
#anova(glm3a)
#car::Anova(glmmtb_all, type = "III")
```

Model 2 summary table
```{r Model 2 table}
knitr::kable(car::Anova(glmmtb_all, type = "III"))
```


```{r Model 2 binomial  checking residuals}
library(DHARMa, quietly = TRUE)
glmtb_resid <- simulateResiduals(glmmtb_all)
testOutliers(glmtb_resid, type = "bootstrap")
plot(glmtb_resid)
#no significant outliers
```


Figure 2
```{r single mix plot for Fig2}

disease2 <- disease2[which(disease2$Date!= "2019.02.01"),]
sum_gen_all_sing_mix2 <- Rmisc::summarySE(disease2, measurevar = c("prop"), groupvars = c("Date3","sing_mix"), na.rm = T)

#summary for Figure 2
sum_gybrk_sing_mix <- Rmisc::summarySE(na.omit(disease2), measurevar = c("prop"), groupvars = c("Date","Date3","sing_mix","Genotype"))

#summary for SFig 3
sumall_gybrk_sing_mix <- Rmisc::summarySE(na.omit(GenRYGBK_sm), measurevar = c("prop"), groupvars = c("Date","sing_mix","Genotype","Coral_stat"))


```

Supplement Figure S3: Proportion of health status by genotypes on both single and mixed frames

```{r Supplemental figure Genotypes by health status}

sumall_gybrk_sing_mix$Date2 <- gsub("[.]", "/", sumall_gybrk_sing_mix$Date)
sumall_gybrk_sing_mix$Date3 <- as.Date(sumall_gybrk_sing_mix$Date2)
                          
sumall_gybrk_sing_mix$Coral_stat <- factor(sumall_gybrk_sing_mix$Coral_stat, levels = c("H","X","D"), labels = c("Healthy","Disease","Dead"))


ggplot(sumall_gybrk_sing_mix[which(sumall_gybrk_sing_mix$Date != "2019.02.01"),], aes(x = Date3, y = prop, group =sing_mix))  + geom_errorbar(aes(ymax = prop+se, ymin = prop-se), position = position_dodge(width = 0.5), color = "gray") + ylab("Proportion of Corals") + theme_bw() + scale_fill_manual(values = c("darkorchid4","gold"), labels = c("mixed", "single"), name = "Diversity")+ theme(text = element_text(size = 16), axis.text.x = element_text(angle = 90),panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "bottom")+ geom_point(aes(fill = sing_mix), pch = 21, color = "black",size = 3, alpha= 0.75, position = position_dodge(width = 0.5)) + xlab("Date") + facet_grid(vars(Genotype), vars(Coral_stat))

```

Individual genotype plots for disease prevalence
```{r Genotype individual plots}
plot.Y <- ggplot(sum_gybrk_sing_mix[which(sum_gybrk_sing_mix$Genotype=="Y"),], aes(x = Date3, y = prop, group =sing_mix))  + geom_errorbar(aes(ymax = prop+se, ymin = prop-se), position = position_dodge(width = 0.5), color = "gray") + ylab("Disease Prevalance") + theme_bw() + scale_fill_manual(values = c("darkorchid4","gold"), labels = c("mixed", "single"), name = "Diversity") + theme(axis.text.x = element_text(angle = 90),panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "bottom")+ geom_point(aes(fill = sing_mix), pch = 21, color = "black",size = 3, alpha= 0.75, position = position_dodge(width = 0.5)) + xlab("Date") + facet_grid(~Genotype) + lims(y = c(0,0.85))

plot.G <- ggplot(sum_gybrk_sing_mix[which(sum_gybrk_sing_mix$Genotype=="G"),], aes(x = Date3, y = prop, group =sing_mix))  + geom_errorbar(aes(ymax = prop+se, ymin = prop-se), position = position_dodge(width = 0.5), color = "gray") + ylab("Disease Prevalance") + theme_bw() + scale_fill_manual(values = c("darkorchid4","gold"), labels = c("mixed", "single"), name = "Diversity") + theme(axis.text.x = element_text(angle = 90),panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "bottom")+ geom_point(aes(fill = sing_mix), pch = 21, color = "black",size = 3, alpha= 0.75, position = position_dodge(width = 0.5)) + xlab("Date") + facet_grid(~Genotype)+ lims(y = c(0,0.85))

plot.R <- ggplot(sum_gybrk_sing_mix[which(sum_gybrk_sing_mix$Genotype=="R"),], aes(x = Date3, y = prop, group =sing_mix))  + geom_errorbar(aes(ymax = prop+se, ymin = prop-se), position = position_dodge(width = 0.5), color = "gray") + ylab("Disease Prevalance") + theme_bw() + scale_fill_manual(values = c("darkorchid4","gold"), labels = c("mixed", "single"), name = "Diversity") + theme(axis.text.x = element_text(angle = 90),panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "bottom")+ geom_point(aes(fill = sing_mix), pch = 21, color = "black",size = 3, alpha= 0.75, position = position_dodge(width = 0.5)) + xlab("Date") + facet_grid(~Genotype)+ lims(y = c(0,0.85))

plot.B <- ggplot(sum_gybrk_sing_mix[which(sum_gybrk_sing_mix$Genotype=="B"),], aes(x = Date3, y = prop, group =sing_mix))  + geom_errorbar(aes(ymax = prop+se, ymin = prop-se), position = position_dodge(width = 0.5), color = "gray") + ylab("Disease Prevalance") + theme_bw() + scale_fill_manual(values = c("darkorchid4","gold"), labels = c("mixed", "single"), name = "Diversity") + theme(axis.text.x = element_text(angle = 90),panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "bottom")+ geom_point(aes(fill = sing_mix), pch = 21, color = "black",size = 3, alpha= 0.75, position = position_dodge(width = 0.5)) + xlab("Date") + facet_grid(~Genotype)+ lims(y = c(0,0.85))

plot.K <- ggplot(sum_gybrk_sing_mix[which(sum_gybrk_sing_mix$Genotype=="K"),], aes(x = Date3, y = prop, group =sing_mix))  + geom_errorbar(aes(ymax = prop+se, ymin = prop-se), position = position_dodge(width = 0.5), color = "gray") + ylab("Disease Prevalance") + theme_bw() + scale_fill_manual(values = c("darkorchid4","gold"), labels = c("mixed", "single"), name = "Diversity") + theme(axis.text.x = element_text(angle = 90),panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "bottom")+ geom_point(aes(fill = sing_mix), pch = 21, color = "black",size = 3, alpha= 0.75, position = position_dodge(width = 0.5)) + xlab("Date") + facet_grid(~Genotype)+ lims(y = c(0,0.85))

```

Figure 2
```{r Figure 2 whole plot}
library(patchwork)

(sing_mix_all + plot_layout(widths = 25))/(plot.G + plot.Y + plot.R + plot.B + plot.K
+ plot_layout(nrow = 1, widths = 25))/plot_layout(guides = "collect") & theme(legend.position = "bottom") & plot_annotation(tag_levels = 'a') 

```


Supplemental Figure 1: Temperature 
```{r temperature}
tempcsv <- read.csv("~/Dropbox/POST DOC/PostDoc Project/Spring 2019/Research/Research/Disease_deep nursery/Microbial Sampling_wholeframes/Temperature/tempcsv.csv", header=FALSE, comment.char="#")

temp <- tempcsv[,2:3]
colnames(temp) <- c("Date_time","temp")
temp <- separate(temp, "Date_time",c("Date","time")," ")
temp$Date <- as.Date(temp$Date, format = "%m/%d/%y")
temp2 <- subset(temp, Date < "2019-08-19")
temp.sum <- summarySE(temp2, measurevar = "temp", groupvars = c("Date"))

martha_r <- read.csv("~/Dropbox/POST DOC/PostDoc Project/Spring 2019/Research/Research/Disease_deep nursery/Microbial Sampling_wholeframes/Temperature/martha_r.csv")

temp3 <-martha_r
colnames(temp3) <- c("Date_time","temp")
temp4 <- separate(temp3, "Date_time",c("Date","time")," ")
str(temp4)

temp4$Date <- as.Date(temp4$Date, format = "%m/%d/%y")

temp5 <- subset(temp4, Date > "2019-05-19" & Date < "2019-08-10")
temp.sum2 <- summarySE(temp5, measurevar = "temp", groupvars = c("Date"))

dim(temp.sum2)
temp.sum2$Loc <- "Martha"
temp.sum$Loc <- "Deep"

temp.both <- data.table(rbind(temp.sum, temp.sum2))
temp.both$tempC <- (temp.both$temp - 32)* (5/9)
templab <- "Temperature (°C)"

templot <- ggplot(temp.both, aes(x = Date, y = tempC)) + geom_point(size = 3, fill = "black", color = "black", pch = 21) + scale_x_date(date_labels = "%m-%d",date_breaks = "10 day" ) + ylab(templab) + theme(text = element_text(size = 16)) + theme_bw() + geom_line()
```




```{r temperature plots}

library(cowplot)
sum_all2_a <- select(sum_all2, Coral_stat, N, prop, se, Date3)
colnames(sum_all2_a)[5] <- "Date"
temp.dis <- merge(sum_all2_a, temp.both, by = "Date")

tempdeadp <- ggplot(temp.dis[which(temp.dis$Coral_stat == "D"),], aes(x = tempC,  y = prop)) + geom_point(size = 3) + theme_bw() + xlab(templab) + ylab("Proportion of Dead Corals") + geom_smooth(se = F, col = "black")

tempdisp <-ggplot(temp.dis[which(temp.dis$Coral_stat == "X"),], aes(x = tempC,  y = prop)) + geom_point(size = 3) + theme_bw() + xlab(templab) + ylab("Proportion of Diseased Corals") + geom_smooth(se = F, col = "gray")
  
healthtemp <- ggplot(temp.dis, aes(x = tempC,  y = prop, group= Coral_stat)) + geom_point(size = 3, aes(color = Coral_stat)) + theme_bw() + xlab(templab) + ylab("Proportion of Corals") + geom_smooth(se = F, aes(color = Coral_stat), method = "lm")+ scale_color_manual(values = c("black","orange", "gray"), name = "Health", labels = c("Dead","Healthy", "Disease"))



plot_grid(templot, healthtemp, labels = "auto")

```



```{r linear model temperature results}
temp.prop.lm <- lm(prop~tempC*Coral_stat, data= temp.dis)
summary(temp.prop.lm)
plot(temp.prop.lm)
car::Anova(temp.prop.lm, type = "III")
coef(temp.prop.lm)
```

```{r linear models by health status}
temp.prop.lmX <- lm(prop~tempC, data= temp.dis[which(temp.dis$Coral_stat == "X"),])
temp.prop.lmD <- lm(prop~tempC, data= temp.dis[which(temp.dis$Coral_stat == "D"),])
temp.prop.lmH <- lm(prop~tempC, data= temp.dis[which(temp.dis$Coral_stat == "H"),])

summary(temp.prop.lmX)
summary(temp.prop.lmD)
summary(temp.prop.lmH)
```

